Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-08-05 alloscore_manuscript/manuscript.Rnw
add Ben's edits to Rnw

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

@@

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
\ref{sec:methods.detailed.integrated_allocation}). IAS was computed for two weight distributions on $K$, one uniform
across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
both scores (Table \ref{tab:multi-k-scores}).

@@

@@

In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. New collaborative work between public health officials and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.

>>>>

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
\begin{align} 
  S_{IAS}(F, y) &= \int S_A(F,y; K) p(K) \, dK \label{eq:IASintegral}\\
    & = \sum_{K=1}^{\infty} S_A(F,y; K) p(K) \quad (\text{assumming non-divisible resource units})\label{eq:IASsum}.
\end{align}
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

\subsubsection{Sensitivity of allocation score to values of K}
\label{sec:sensitivity}

@@

The integrated allocation score (IAS) summarizes allocation scores (AS) across a range of possible values of the
constraint ($K$), allowing one to assign higher weights to more likely values of $K$ (Section
\ref{sec:methods.detailed.integrated_allocation}). 
Calculating exact IAS as defined in equation \eqref{eq:IASsum} is computationally intensive so an approximate IAS, denoted $\hat{S}(F,y)$, was used such that
\begin{align*}
	\hat{S}(F,y) = \sum_{k=1}^{600} S_A(F,y; 200k) p(200k),
\end{align*}
where summation terms were calculated at every 200 beds with an upper bound of 60,000, or four times the estimated available beds.
An approximate IAS was computed for two weight distributions on $K$, one uniform
across the entire range and the other a truncated normal distribution centered at 15,000 (the orange and blue shaded areas in Figure
\ref{fig:multi-k}A, respectively). Both versions of the IAS were correlated with the original AS for $K=15,000$, with
the higher correlation coming from the weighting centered at $K=15,000$ (Figure \ref{fig:multi-k}B). Model rankings
based on the AS and the centered IAS were roughly similar, with the top and bottom three approaches being the same for
both scores (Table \ref{tab:multi-k-scores}).

@@

Additionally, during an epidemic, public health decision makers may be faced with an ongoing series of decisions over time in response to changes in conditions such as resource availability or the current effective allocations of resources. The AS presented here looks at a single snapshot in time  and ignores many of the possible complexities related to adjusting existing allocations. The example presented in section 
\ref{sec:application} is based on a prospective estimate of bed availability, but one could also use observed total resource availability to target an optimal set of prospective AS parameters. 
Alternatively, the impact of unmet need often differs across different locations and demographics, which could be accounted for by incorporating a function of location and demographics which defines the cost of unmet need. Future work is needed to explore the use of AS while accounting for the dynamic nature of disease spread and resource availability.

@@

In conclusion, we argue that when possible, the way modelers and policymakers view and evaluate forecasts should more explicitly depend on
the specific decision-making context. Defaulting to standard forecast evaluation metrics can mask the utility (or
disutility) of certain forecasts, or lead to forecasts being used in decision making contexts very different from those
for which they are able to offer useful guidance. New collaborative work between decision-makers and modeling teams is
needed to assess the value and relevance of the initial findings presented here, including real-time pilot studies or
simulation exercises that could be used to inform further development of new or alternative scoring metrics. We see this
work as an initial overture for what we hope will grow to be a large, collaborative body of work more closely coupling
applied epidemiological forecasting with public health decision making.
